Ecological and evolutionary implications of starvation and body size

### 2016-07-25 Manuscript/2.tex
Kempes edits to v2

\vspace{2mm}
\noindent {\bf Role of allometry} \\ 
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
The parameters in the NSM cannot be freely navigated in biologically-reasonable settings, and our first challenge is to constrain the covariation of rates in a principled and biologically meaningful manner. % we have not described the realistic regimes occupied by organisms where the
Allometric scaling relationships highlight common constraints and average trends across large ranges in body size and species diversity. Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model (cf. \cite{Yodzis:1992hg}). We are able to define the regime of dynamics occupied by the entire class of mammals along with the key differences between the largest and smallest mammals.

%Second paragraph
Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, and thus can be estimated based on known allometric constraints. 
The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{West:2002it} and plays a central role in a variety of scaling relationships.
Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{DeLong:2010dy}. Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{Kempes:2012hy}. More specifically, the interspecific trends in growth rate can be approximated by $\lambda = \lambda_0 M^{\eta-1}$. This relationship is derived from the simple balance
\begin{eqnarray}
\label{balance}
B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m
\end{eqnarray}
 [a and b notation --- these parameters are easily measured bioenergetic parameters which are often approximately invariant across organisms of vastly different size. Our notation seeks to illustrate that the allometric model fundamentally depends on a small number of free parameters.]
 where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is the metabolic rate to support an existing unit of mass, and $m$ is the mass at any point in development. It is useful to explicitly write this balance because it can also be modified to understand the rates of both starvation and recovery from starvation. [Spell out the connection to nutritional state more explicitly] [As we will see it is possible to derive both sigma and rho from this balance]

For the rate of starvation, we make the simple assumption that an organism must meet its maintenance requirements using digested mass as the sole energy source. This assumption implies the simple metabolic balance 
\begin{eqnarray}
\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
\end{eqnarray}
where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which may differ from  $E_{m}$, the energy required to synthesis a unit of biomass. Give the adult mass, $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
\begin{eqnarray}
m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
\end{eqnarray}
Considering that only certain tissues can be digested for energy, for example the brain cannot be degraded to fuel metabolism, we define the rate for starvation and death by the timescales required to reach specific fractions of normal adult mass. We define $m_{starve}=\epsilon M$ where it could be the case that organisms have a systematic size-dependent requirement for essential tissues, such as the minimal bone or brain mass. For example, considering the observation that body fat in mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$, and assuming that once this mass is fully digested the organism begins to starve, would imply that $\epsilon=1-f_{0}M^{\gamma}/M$. Taken together the time scale for starvation is given by
\begin{equation}\label{eq:sigma}
t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
\end{equation}
The starvation rate is $\sigma=1/t_{\sigma}$, which implies that $\sigma$ is independent of adult mass if $\epsilon$ is a constant, and if $\epsilon$ does scale with mass, then $\sigma$ will have a factor of $1/\log\left(1-f_{0}M^{\gamma}/M\right)$. In either case $\sigma$ does not have a simple scaling with $\lambda$ which is important for the dynamics that we later discuss.

@@

The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass. The general solution to Equation \ref{balance} is given by
\begin{eqnarray}
m\left(t\right)=c\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
\end{eqnarray}
with $a=B_{0}/E_{m}$, $b=B_{m}/E_{m}$, and $c=(a/b)^{1/(\eta-1)}$. We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of 
\begin{eqnarray}
t_{\rho}=\frac{\log \left(1-\left(cM \right)^{1-\eta }\right)-\log \left(1-\left(c\epsilon M \right)^{1-\eta }\right)}{(\eta -1) b}.
\end{eqnarray}
%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
%where $c=(a/b)^{1/(\eta-1)}$.
Although these rate equations are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
Investigating other classes of organisms requires only substituting the energetic and scale parameters shown in Table \ref{param}.
Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species. 
%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.

@@

%Bifurcation space as a function of mass
As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, and the bifurcation space shown in Fig. \ref{fig:fp} cannot be freely navigated if assuming biologically reasonable parameterizations.
Given the parameterization for terrestrial endotherms shown in Table \ref{param} with mass $M$ as a free parameter, we show that $\sigma$ and $\rho$ are constrained to a small window of potential values (Fig. \ref{fig:hopf}), thus confining dynamics to the steady state regime for all realized body size classes.
%, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig. \ref{fig:hopf}) results in little qualitative difference.
This suggests that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than mammals with larger body size. %, and
%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.

Allometric constraints have been invoked to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, such that period $\propto M^{0.25}$, however this relationship seems to hold only for some species \cite{Hendriks:2012fc} and competing explanations exist \cite{Kendall:1999iy,Hogstedt:2005cr}.
%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic \cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
Statistically significant support for the existence of population cycles among mammals is predominantly based on time-series for smaller bodied mammals \cite{Kendall:1998hl}, though we acknowledge that longer generation times precludes similar quality data for larger organisms.
%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
We thus obtain a specific prediction from our model: population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting. %, emphasizing the role of starvation on dynamics.

%Extinction to the left and the right
Higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq. \ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq. \ref{eq:ss}). %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
By allowing the rate of starvation to vary, we assessed extinction risk across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig. \ref{fig:fp}.
As expected, higher rates of extinction correlated with both low and high values of $\sigma/\rho$; for low values the higher extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig. \ref{fig:ext}).
For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively constrained range of $\sigma/\rho$, extinction probabilities are minimized.

As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refuge, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.
Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
Nevertheless, our finding that allometrically-determined energetic rates place the system within this low extinction probability region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.

@@

%Minimal mass bound
The NSM correctly predicts that species with smaller masses have larger steady state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig. \ref{fig:mass}a,b).
Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases, while the rate of recovery declines super-exponentially.
This decline in $\rho$ occurs when body fat percentage is $1 - 1/(cM) \approx 2\%$, whereupon consumers have no eligible route out of starvation.
Compellingly, this dynamic bound determined by the rate of energetic recovery is close to the minimum observed mammalian body size of ca. 1.3-2.5 grams (Fig. \ref{fig:mass}b,c), a range that occurs as the recovery rate begins its decline.
In addition to known transport limitations \cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.

%Maximum mass bound
Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
Instead, we examine a potential upper bound to body mass by assessing population invasibility with respect to a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we can assess invasibility as a function of organismal mass by determining which consumer steady state is larger over $\chi$.
We find that for $1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher steady state density for the invader population ($H^{\prime *}+F^{\prime *}>H^*+F^*$), such that it has an intrinsic advantage over the resident population.
For $M>10^6$, however, there is an increasing range of $\chi < 0$ such that leaner individuals have the advantage, and this is due to the changing covariance between energetic rates as a function of modified energetic reserves.

@@

\begin{figure*}
\centering
\includegraphics[width=1\textwidth]{fig_Invasion.pdf}
\caption{ % 58 words
}
\label{fig:invasion}
\end{figure*}

>>>>

\vspace{2mm}
\noindent {\bf Role of allometry} \\ 
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
The NSM describes a broad range of dynamics, yet organisms are likely unable to access most of the total parameter space. Here we use allometric scaling relationships to constrain the covariation of rates in a principled and biologically meaningful manner. % we have not described the realistic regimes occupied by organisms where the
Allometric scaling relationships highlight common constraints and average trends across large ranges in body size and species diversity. Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model (cf. \cite{Yodzis:1992hg}). We are able to define the regime of dynamics occupied by the entire class of mammals along with the key differences between the largest and smallest mammals.

%Second paragraph
Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, which can be used to describe a variety of organism features \cite{}. The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{West:2002it} and scales as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to vary around $2/3$ or $3/4$ for metazoans \cite{}, and has taxonomic shifts for unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{DeLong:2010dy,Kempes:2012hy}. Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{Kempes:2012hy,more}. More specifically, the interspecific trends in growth rate can be approximated by 
\begin{eqnarray}
\lambda = \lambda_0 M^{\eta-1}.
\label{lambda}
\end{eqnarray}
This relationship is derived from the simple balance
\begin{eqnarray}
\label{balance}
B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m
\end{eqnarray}
where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is the metabolic rate to support an existing unit of mass, and $m$ is the mass at any point in development. It is useful to explicitly write this balance because it can also be modified to understand the timescales of both starvation and recovery from starvation as we show below.

For the rate of starvation, $\sigma$, we are interested in the time that it takes an organism to go from healthy adulthood to a smaller size where reproduction is impossible. This process is defined by the energy balance in Equation \ref{balance} where we make the simple assumption that an organism must meet its maintenance requirements using the digestion of existing mass as the sole energy source. This assumption implies the simple metabolic balance 
\begin{eqnarray}
\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
\end{eqnarray}
where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which differs from $E_{m}$ \cite{}, the energy required to synthesis a unit of biomass. Given the adult mass, $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
\begin{eqnarray}
m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
\end{eqnarray}
Considering that only certain tissues can be digested for energy, for example the brain cannot be degraded to fuel metabolism, we define the rate for starvation and death by the timescales required to reach specific fractions of normal adult mass. We define $m_{starve}=\epsilon M$, where $\epsilon<1$ is the fraction of adult mass where reproduction ceases. This fraction will change if tissue composition systematically scales with adult mass. For example, considering the observation that body fat in mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$, and assuming that once this mass is fully digested the organism begins to starve, would imply that $\epsilon=1-f_{0}M^{\gamma}/M$. Taken together the time scale for starvation is given by
\begin{equation}
\label{eq:sigma}
t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
\end{equation}
The starvation rate is then $\sigma=1/t_{\sigma}$, which will scale with adult mass following $1/\log\left(1-f_{0}M^{\gamma}/M\right)$. In either case $\sigma$ does not have a simple scaling with $\lambda$ (Equation \ref{lambda}) which is important for the dynamics that we later discuss.

@@

The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass. The solution to Equation \ref{balance} is given by
\begin{eqnarray}
m\left(t\right)=\left(\frac{B_{0}}{B_{m}}\right)^{1/(\eta-1)}\left[1-\left(1-\frac{B_{m}}{B_{0}}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
\end{eqnarray}
We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
\begin{eqnarray}
t_{\rho}=\frac{E_{m}\left(\log \left(1-\frac{B_{0}}{B_{m}}\left(M \right)^{1-\eta }\right)-\log \left(1-\frac{B_{0}}{B_{m}}\left(\epsilon M \right)^{1-\eta }\right)\right)}{(\eta -1) B_{m}}.
\end{eqnarray}
%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
%where $c=(a/b)^{1/(\eta-1)}$.
Although these rate equations are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$. Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species. 
%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.

@@

%Bifurcation space as a function of mass
As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, and the bifurcation space shown in Fig. \ref{fig:fp} is navigated via covarying parameters. Given the parameterization for terrestrial endotherms we find that $\sigma$ and $\rho$ are constrained to a small window of potential values (Fig. \ref{fig:hopf}) for the known range of body sizes $M$. We find that the dynamics for all mammalian body sizes is confined to the steady-state regime which does not contain limit cycles.
%, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig. \ref{fig:hopf}) results in little qualitative difference. These results suggest that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than mammals with larger body size. Our model predicts that population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.%, and
%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.

Previous studies have used allometric constraints explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$, however this relationship seems to hold only for some species \cite{Hendriks:2012fc} and competing explanations [related to ....] exist \cite{Kendall:1999iy,Hogstedt:2005cr}. 
%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic \cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
Statistically significant support for the existence of population cycles among mammals is predominantly based on time-series for smaller bodied mammals \cite{Kendall:1998hl}, where we our model would predict much longer and more pronounced transient dynamics given how close these points are to the Hopf bifurcation. We acknowledge that longer generation times, and more difficult measurement procedures, precludes similar quality data for larger organisms.
%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
 %, emphasizing the role of starvation on dynamics.

%Extinction to the left and the right
\paragraph{Extinction risk}
Within our model higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality. However, from the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}). We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq. \ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq. \ref{eq:ss}). %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
By allowing the rate of starvation to vary, we assessed extinction risk across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig. \ref{fig:fp}.
As expected, higher rates of extinction correlated with both low and high values of $\sigma/\rho$; for low values the higher extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig. \ref{fig:ext}).
For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively constrained range of $\sigma/\rho$, extinction probabilities are minimized.

We find that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refuge. These values are close enough to the Hopf bifurcation to avoid low steady state densities, and far enough away to avoid large-amplitude transient cycles. The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that avoid extinction, and this could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
Nevertheless, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive -- and constrain -- natural animal populations.

@@

%Minimal mass bound
A lower bound on mammalian body size is given by $\epsilon=1$, where mammals have no reserves and immediately starve, which occurs at a size of $M=value$. This calculation gives an extreme limit on size but does not account for the subtleties of starvation dynamics that may limit body size. The NSM correctly predicts that species with smaller masses have larger steady-state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig. \ref{fig:mass}a,b).
Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases, while the rate of recovery declines super-exponentially.
This decline in $\rho$ occurs when body fat percentage is $1 - 1/(\left(B_{0}/B_{m}\right)^{1/\left(\eta-1\right)}M) \approx 2\%$, whereupon consumers have no eligible route out of starvation.
Compellingly, this dynamic bound determined by the rate of energetic recovery is close to the minimum observed mammalian body size of ca. 1.3-2.5 grams (Fig. \ref{fig:mass}b,c), a range that occurs as the recovery rate begins its decline.
In addition to known transport limitations \cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation. This work mirrors other efforts where coincident limitations seem to limit the smallest possibilities for life within a particular class or organisms \cite{kempes}.

%Maximum mass bound
%Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
We examine a potential upper bound to body mass by assessing population invasibility with respect to a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we assess invasibility, as a function of organismal mass, by determining which consumer has a higher steady state given $\chi$.
We find that for $1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher steady state density for the invader population ($H^{\prime *}+F^{\prime *}>H^*+F^*$), such that it has an intrinsic advantage over the resident population.
However, for $M>10^6$, leaner individuals ($\chi < 0$) have an advantage, and this is due to the changing covariance between energetic rates as a function of modified energetic reserves.

@@

\begin{figure*}
\centering
\includegraphics[width=0.5\textwidth]{fig_Invasion.pdf}
\caption{ % 58 words
}
\label{fig:invasion}
\end{figure*}
